Developing a preprint culture in biology

### 2012-09-26 ms.tex
first round of editing for the first two sections

\begin{abstract}

\end{abstract}

@@

Public preprints servers allow authors to make their manuscripts publicly
available before, or in parallel to, submitting them to journals for traditional
peer-review. This idea gained popularity 20 years ago with the advent of arXiv,
an open preprint server widely used in the physical sciences \cite{gin11}, but
it also part of other fields' culture. Paul Krugman noted that, in economics,
the \emph{traditional model of submit, get refereed, publish, and then people
will read your work broke down a long time ago. In fact, it had more or less
fallen apart by the early 80s} \cite{kru12}. In addition to a section in arXiv,
economists have also the RePEc (Research Papers in Economics) initiative, which
aims to create an archive of working papers, manuscripts, and book chapters.

The idea behind preprint servers is fundamentally simple: to make the results
available to the scientific community as soon as possible rather than wait until
the peer-review process is fully completed. The point of arXiv and open
preprints servers is not to avoid the peer-review process. Almost all
manuscripts on arXiv are submitted to peer-review.  Sharing drafts of research
papers on preprint servers has numerous advantages. These include rapid
dissemination of work in progress to a wider audience that includes not only
scientists and the general public, but also readers in developing countries
where access to subscription journals are often a limiting factor. Further,
preprints provide the opportunity to solicit feedback from a larger pool of
reviewers. As a consequence, preprint servers increase the number of
opportunities for review and revision prior to publication, resulting in higher
quality submissions which could also alleviate reviewer burden.

@@

The first and most often discussed advantage of arXiv and open preprints is
speed \ref{fig:map}. The time between submission and the official publication of
a manuscript can be measured in months, sometime in years. For all this time,
the research is only known to a select few: colleagues, editors, reviewers.
Thus, the science cannot be used, discussed, or reviewed by the wider scientific
community. It is a problem for both scientists and publishers. Manuscripts that
are unknown cannot be integrated and thus take more time to be cited.

The review process as a whole is critically over-loaded, because the number of
active scientists increases, because the pressure to publish increases, and
because of an effect dubbed ``the tragedy of the reviewers commons'' REF.  In
the same times, rejection rates are high in most journals (REF), and when the
not invited to submit a revisions, authors are left with the impression that
they must start the whole process all over again. It's thus no surprise that
different initiatives emerged over the last few years, to decrease the time
spent in review. XXX et coll. (REF) called for the recycling and reuse of
peer-reviews: by attaching previous reviews, and detailed replies, to a new
submission, both the editor and the referees can gauge the work done on the
manuscript, and perhaps evaluate it with less prejudice. In a similar way, the
\emph{Peerage of Science} initiative allows authors to seek anonymous pre-
review by their peers. Some journals (LIST?) now accept to publish papers
which received good evaluations, effectively outsourcing the review process. A
widespread use of preprint servers can achieve the same goal of reducing the
time spent in review. By putting a manuscript out there for open comments and
criticisms, the authors will receive valuable feedback, and can improve the
version which will be submitted. With a rich enough community of scientists
depositing preprints, and commenting on them, the process of an open pre-
review can become widespread, and will overall increase the quality of first
submissions.

\begin{figure}[ht!] \centering\includegraphics[width=0.50\textwidth]{map.pdf}
\caption { It can take several months, and even a few years, before a submitted
paper is officially published and citable. During this time, few people are
aware of the research that has been done (typically, close colleagues are
given access to the preprints). With public preprint servers, the science is
immediately available and can be openly discussed, analysed, and integrated
into current research. It benefits both science and publishers. Both want the
papers to be well-known and cited, and public preprints make it possible to
integrate research even before publication, greatly improving immediacy.  }
\label{fig:map}
\end{figure}

Preprint servers also establish priority in a fair way.  Some manuscripts will
spend much more time in the review process than others.  Public preprint servers
offer a much fairer way to establish intellectual priority by making the work
available when done, even if the exact organisation of the manuscript may
change. Surprisingly, there is a perception in biology that public preprints
make it easier to steal ideas, as if scientific ideas only took form in
published material.  Mathematicians and physicists have embraced arXiv in part
to establish priority in a fair way\cite{cal12}.

Some of the responses to public preprints are surprising since they are,
essentially, the same as exchanging preprints among colleagues.  Prepublication
reviews by a small network of colleagues is an important part of the scientific
process, which is attested by the fact that nearly all published papers
acknowledge comments by people not listed as co-authors.  Preprint servers
simply offer a way to extend this network of colleagues to the entire scientific
community. It ensures that science is not constrained by small networks of
scientists exchanging ideas.  Paul Ginsparg created arXiv.org in part for
democratic reasons: he wanted everyone from students in small universities to
Ivy-League professors to have access to the most recent scientific \emph{ideas}.
Ginsparg's revolutionary idea was simply to use the power of the internet for
preprints, not just for the end product, so the process can be open from A to Z,
instead of being just open at the end of the process.
% TP One important point to make: the MS and the idea are different things.  We
% should focus more on the later...

@@

While submitting to public preprints servers is still uncommon in biology,
preprints are becoming more common in biological sciences. The quantitative
biology section in arXiv is experiencing faster growth in submissions than any
other field \cite{cal12}. Also, most scientific journals are preprint-friendly:
Nature, PLOS, BMC, PNAS, Science (mostly) \ref{table:policies}, and all the
journals from Elsevier and Springer. Very recently, the Ecological Society of
America recently changed its policy to allow public preprints (REF). In our
field, few scientific publications will not consider a manuscript submitted to
arXiv.  Still, many journals adopt a ``by default'' hostile attitude towards
preprints, mostly due to the lack of clear policy of the publishers. As an
example, Wiley-Blackwell, which publishes some of the leading journal in the
field, has no official policy on the subject \ref{table:policies}.

\begin{table*}
    \centering
    \begin{tabular}{|ll|}
    \hline
    Publisher                                   & Policy \\
    \hline
    Springer                            	& Accept \\
    BMC                                 	& Accept \\
    Elsevier                            	& Accept \\
    Nature Publishing Group             	& Accept \\
    Public Library of Science           	& Accept \\
    Royal Society                       	& Accept \\
    National Academy of Science (USA)           & Accept \\
    Ecological Society of America       	& Accept \\
    Oxford Journals                             & Accept \\
    Science                             	& Ambiguous \\
    Wiley-Blackwell                       	& No general policy \\
    British Ecological Society                  & No answer to our query \\
    \hline
    \end{tabular}
    \caption{Policies for important publishers in biology. Some publishers
tolerate preprints except for a few of their medical journals, e.g.: Journal
of the National Cancer Institute from Oxford and The Lancer from Elsevier.}
    \label{table:policies}
\end{table*}

There are other ``cultural'' barriers to adoption of preprint servers by
biologists.  The most notable preprint server,
arXiv, is designed to work best with \TeX{}, a nearly universal document
preparation system among physicists and mathematicians.  Jackson
\cite{jackson2002preprints} further argues that \TeX{} introduced an ``open
source mindset'' to its users, who now freely share their research findings as
well as their software.  Many biologists instead use proprietary software to
prepare their research findings, and many journals officially prefer Microsoft
Word documents as submissions.  The recent discipline-wide adoption of free
software packages such as R and associated document generation systems
\cite{xie2012}, as well as interest in open access publishing, has coincided
with recent rise in use of preprint servers by biologists, supporting Jackson's
claims. \cite{xie12}

@@

% Rough outline:
% 1. Description of arXiv (origins, funding, maintenance)
% 2. Submission process
% 3. Role within math and physics community (nearly universal), including recent changes
% 4. Nature of citations and "priority"
%
% My biggest problem here is how to cite my sources: the information
% on funding comes from a press release from Cornell University
% Libraries; other basic "raw data" comes straight from the arXiv
% webpages. How do I cite that?
% -- Joel

@@

Figshare (\href{http://figshare.com/}{http://figshare.com/}) is an open server
that allows scientists to submit any research output: manuscript, figures,
datasets, videos, theses, presentations, and so on. There are no rules to limit
what constitutes a research output and, unlike arXiv, there is no endorser
system (although the administrator reserve the right to remove inappropriate
content). All figshare content (article, figures, datasets) have a unique
digital object identifier (DOI) like any journal article. Tags are used to
classify the content and new tags can be made. All content can be commented and
is licensed under the Creative Commons (CC-BY) license, except datasets which
are published under CC0.

@@

% Written from memory: to verify!

@@

@@

% Short list of disadvantages

@@

%The average time to publication vary a lot between journal and can be as low as
%104 days (Evolution for 2011) to 213 (PLOS One in 2010)

>>>>

% No abstract needed.

@@

Public preprints servers allow authors to make their manuscripts publicly
available before, or in parallel to, submitting them to journals for traditional
peer-review. This idea gained popularity 20 years ago with the advent of arXiv,
an open preprint server widely used in physics and mathematics \cite{gin11}. 
It is also part of other fields' culture. Paul Krugman noted that, in economics,
the \emph{traditional model of submit, get refereed, publish, and then people
will read your work broke down a long time ago. In fact, it had more or less
fallen apart by the early 80s} \cite{kru12}. In addition to a section in arXiv,
economists have also the RePEc (Research Papers in Economics) initiative, which
aims to create an archive of working papers, manuscripts, and book chapters.

The idea behind preprint servers is fundamentally simple: to make the results
available to the scientific community as soon as possible rather than wait until
the peer-review process is fully completed. The point of arXiv and open
preprints servers is not to avoid the peer-review process. Almost all
manuscripts on arXiv are submitted to peer-review.  The point is that sharing
ideas as quickly as possible using preprint servers has numerous advantages.
These include rapid dissemination of work in progress to a wider audience, not
only scientists and the general public but also readers in developing countries
where access to subscription journals are often a limiting factor. Further,
preprints provide the opportunity to solicit feedback from a larger pool of
reviewers. As a consequence, preprint servers increase the number of
opportunities for review and revision prior to publication, resulting in higher
quality submissions which could also alleviate reviewer burden.

@@

The first and most often discussed advantage of arXiv and open preprints is
speed \ref{fig:map}. The time between submission and the official publication of
a manuscript can be measured in months, sometime in years. For all this time,
the research is only known to a select few: colleagues, editors, reviewers.
Thus, the science cannot be used, discussed, or reviewed by the wider scientific
community. It is a problem for both scientists and publishers. Manuscripts that
are unknown cannot be integrated and thus take more time to be cited. It has
been shown that high-energy physics, with its high arXiv submission rate, had
the highest immediacy among physics and mathematics \cite{pra05}.

The review process as a whole is critically over-loaded, because the number of
active scientists increases, because the pressure to publish increases, and
because of an effect dubbed ``the tragedy of the reviewers commons'' REF.  At
the same times, rejection rates are high in most journals (REF), and when not
invited to submit a revisions, authors must start the whole process all over
again. It's thus no surprise that different initiatives emerged over the last
few years, to decrease the time spent in review. XXX et coll. (REF) called for
the recycling and reuse of peer-reviews: by attaching previous reviews, and
detailed replies, to a new submission, both the editor and the referees can
gauge the work done on the manuscript, and perhaps evaluate it with less
prejudice. In a similar way, the \emph{Peerage of Science} initiative allows
authors to seek anonymous pre-review by their peers. Some journals now accept
to publish papers which received good evaluations, effectively outsourcing the
review process. A widespread use of preprint servers can achieve the same goal
of reducing the time spent in review. By putting a manuscript for open
comments and criticisms, the authors will receive valuable feedback and can
improve the version which will be submitted. With a rich enough community of
scientists depositing preprints, and commenting on them, the process of an open
pre-review can become widespread and will overall increase the quality of
first submissions.

\begin{figure}[ht!] \centering\includegraphics[width=0.50\textwidth]{map.pdf}
\caption { It can take several months, and even a few years, before a submitted
paper is officially published and citable.  The average time to publication
varies greatly between journals and can be as low as 104 days (Evolution for
2011) to 213 (PLOS One in 2010).  Meanwhile, few people are aware of the
research that has been done since, typically, only close colleagues are given
access to the preprints. With public preprint servers, the science is
immediately available and can be openly discussed, analysed, and integrated into
current research. It benefits both science and publishers. Both want the papers
to be well-known and cited, and public preprints make it possible to integrate
research even before publication, greatly improving immediacy.  }
\label{fig:map} \end{figure}

Public preprint servers offer a much fairer way to establish intellectual
priority by making the work available when done. Some manuscripts will spend
much more time in the review process than others. Surprisingly, there is a
perception in biology that public preprints make it easier to steal ideas, as if
a scientific idea was only complete once published in a peer-reviewed journal.
Mathematicians and physicists have embraced arXiv in part to establish priority
in a fair way and biologists should do so too, or at least not be afraid to
submit to make their research public for fear of been snooped \cite{cal12}.

Some of the responses to public preprints are surprising since they are,
essentially, the same as exchanging preprints among colleagues.  Prepublication
reviews by a small network of colleagues is an important part of the scientific
process, which is attested by the fact that nearly all published papers
acknowledge comments by people not listed as co-authors.  Preprint servers
simply offer a way to extend this network of colleagues to the entire scientific
community. It ensures that science is not constrained by small networks of
scientists exchanging ideas.  Paul Ginsparg created arXiv.org in part for
democratic reasons: he wanted everyone from students in small universities to
Ivy-League professors to have access to the most recent scientific \emph{ideas}.
Ginsparg's revolutionary idea was simply to use the power of the internet for
preprints, not just for the end product, so the process can be open from A to Z,
instead of being just open at the end of the process.

@@

While submitting to public preprints servers is still uncommon in biology,
preprints are becoming more common in biological sciences. The quantitative
biology section in arXiv is experiencing faster growth in submissions than any
other field \cite{cal12}. Also, most scientific journals are preprint-friendly:
Nature, PLOS, BMC, PNAS, Science (mostly) \ref{table:policies}, and all the
journals from Elsevier and Springer. Very recently, the Ecological Society of
America and the Genetics Society of America changed their policies to allow
public preprints. Few scientific publications will not consider a manuscript
submitted to a public preprint server.  Still, a few journals adopt a ``by
default'' hostile attitude towards preprints, mostly due to the lack of clear
policy of the publishers, or perhaps because a preprint culture has not
developed in biology. As an example, Wiley-Blackwell, which publishes some of
the leading journal in the field, has no official policy on the subject
\ref{table:policies}.

\begin{table*}
    \centering
    \begin{tabular}{|ll|}
    \hline
    Publisher                                   & Policy \\
    \hline
    Springer                            	& Accept \\
    BMC                                 	& Accept \\
    Elsevier                            	& Accept \\
    Nature Publishing Group             	& Accept \\
    Public Library of Science           	& Accept \\
    Genetics Society of America                 & Accept \\
    Royal Society                       	& Accept \\
    National Academy of Science (USA)           & Accept \\
    Ecological Society of America       	& Accept \\
    Oxford Journals                             & Accept \\
    Science                             	& Ambiguous \\
    Wiley-Blackwell                       	& No general policy \\
    British Ecological Society                  & No answer to our query \\
    \hline
    \end{tabular}
    \caption{Policies for important publishers in biology. Some publishers
tolerate preprints except for a few of their medical journals, e.g.: Journal
of the National Cancer Institute from Oxford and The Lancer from Elsevier.}
    \label{table:policies}
\end{table*}

There are other barriers to adoption of preprint servers by biologists.  The
most notable preprint server, arXiv, is designed to work best with \LaTeX, the
dominant document preparation system among physicists and mathematicians.
Jackson \cite{jackson2002preprints} further argues that \LaTeX introduced an
open source mindset to its users, who now freely share their research
findings as well as their software.  Many biologists instead use proprietary
software to prepare their research findings, and many journals officially prefer
Microsoft Word documents as submissions.  The recent discipline-wide adoption of
free software packages such as R and associated document generation systems
\cite{xie2012}, as well as interest in open access publishing, has coincided
with recent rise in use of preprint servers by biologists, supporting Jackson's
claims \cite{xie12}.

@@

@@

Figshare (\href{http://figshare.com/}{http://figshare.com/}) is an open server
that allows scientists to submit any research output: manuscript, figures,
datasets, videos, theses, presentations, and so on. There are no rules to limit
what constitutes a research output and, unlike arXiv, there is no endorser
system. All figshare content has a unique digital object identifier (DOI) like
any journal article, thus offering a permanent and stable link to the content.
A flexible tag system is used to classify the content. All content can
be commented and is licensed under the Creative Commons (CC-BY) license, except
datasets which are published under CC0.

One of the biggest advantage of figshare over arXiv is that is it not limited to
quantitative sciences. arXiv.org has sections on quantitative biology but might
not be appropriate for non-quantitative work. With its flexible approach to
preprints, figshare offer an important alternative to arXiv for empirical
biologists. Furthermore, by allowing all

@@

@@

\subsection{Github et al.?}

This manuscript was developed entirely as an open project on github. github is a
hosting service for projects under the git revision control tool. git is a
decentralized revision control system used primarily to develop software. It
allows the users to share code, track changes, and merge versions. For example,
during the development of this manuscript, one author would fork the project
(that is: make a personal copy), modify it, and then the changes were merged
back into the main branch. These techniques are widely used in open source
software development. It goes even farther than submitting preprints by opening
the entire writing process.

@@

\section{Funding}

PDP is funded by an Alexander Graham Bell scholarship from the National Sciences
and Engineering Council of Canada.

DG is funded by a Discovery Grand from the National Sciences and Engineering
Council of Canada and by the Canada Research Chair program.

\section{Acknowledgements}

@@
